\chapter{Concluding Remarks}
\label{chap:Concl}

After giving an overview of related web technologies and the research
related to the academic fields of information visualizations and
responsive visualizations, this thesis introduced RespVis, a new
open-source software library to create responsive visualizations for
the web.  RespVis has been designed as an extension of the D3 library
and renders visualizations as SVG documents styled with CSS.  The most
novel contribution of this work is a custom layouter that uses the
browser's own layout engine to enable visualization authors to
configure the layout of SVG-based visualization components via CSS.
Since rearranging content is one of the main techniques of responsive
web design, enabling visualization authors to use CSS layout
mechanisms like Flexbox and Grid to reposition visualization
components leads to much better responsive capabilities than merely
allowing them to change their styles.  Relying on CSS for a large
amount of a visualization's configuration also leads to visualization
authors benefitting from being able to utilize CSS media queries for
responsive styling and from the simplicity of using a tool they are
already familiar with.  Furthermore, since RespVis' API is mostly
meant for configuring a visualization's content and behavior, it can
be much more minimal than it would be if the complete style of a
visualization would also be configured via it.  This minimal API and
RespVis' reliance on standards like SVG and CSS for rendering and
configuring its visualizations result in much less likelihood that
visualization authors are limited by API restrictions.  Due to all of
these reasons, it is evident that RespVis has the potential to be a
very effective library for the creation of responsive visualizations
after some more improvements are made to it.
