\chapter{Information Visualization}

\label{chap:InfoVis}

\TODO{Explain InfoVis}

\TODO{InfoVis vs. SciVis vs. GeoVis vs. DataVis}

\TODO{Mention purposes of InfoVis}
\TODO{Mention Anscombe's Quartet}
\TODO{Mention human visual perception}
\TODO{Mention preattentive processing}

\TODO{Explain why interactivity is such an important concept in visualization}

\TODO{Mention the history of information visualization}
\TODO{Supposedly, there is a good book on this topic. Find it and cite it. History of data visualization or something like that.}
\TODO{Add large image with timeline and visualizations}

\section{Responsive Web Design}

\TODO{Define responsive web design} \TODO{Mention adapting to characteristics of consuming device} \TODO{Mention text wrapping} \TODO{Mention }

\TODO{Mention importance of responsive web design} \TODO{Add chart showing the rise of mobile devices} 

\section{Responsive Information Visualization}

\TODO{Mention necessity of responsive visualizations in context of responsive web design} \TODO{Mention visualizations being important blocks of content that can not be ignored when designing responsive web pages}

\TODO{Mention scalable vs responsive visualizations} \TODO{Mention responsive layout} \TODO{Mention responsive display density} \TODO{Mention responsive interaction}

\TODO{Mention the book Building responsive data visualizations for the web} \TODO{Mention focus on scalable visualizations rather than responsive visualizations}

\subsection{Responsive Visualization Patterns}

\TODO{Mention patterns in software development}
\TODO{Mention patterns in responsive web design}
\TODO{Mention patterns in visualizations}

\TODO{Mention patterns defined by Adler et. al}
\TODO{
    1. Rotating axis labels
    2. Removing axis ticks
    3. Shortening strings
    4. Flipping visualizations by 90 degrees
    5. Repositioning visual components
    6. Zooming
    7. Filtering
    8. Reducing data density
    9. Changing chart types
}

\TODO{Mention high-level actions for responsive adaptations of visualization as defined by Hoffswell et. all}
\TODO{Identified by analyzing 231 visualizations from various sources}
\TODO{
    1. No change
    2. Resizing visualizations
    3. Repositioning visual components
    4. Adding visual components
    5. Modifying visual components
    6. Removing visual components
}

\TODO{Mention patterns as defined by Kim et. al (2021 paper)}

\subsection{Responsive Visualization Examples}

\section{Information Visualization Libraries}

\subsection{Chartist}

\subsection{Highcharts}

\subsection{ECharts}

\subsection{...?}

\subsection{D3}

\TODO{Mention that D3 is successor of Protovis}

\subsection{D3FC}

\TODO{Mention the mixed HTML/SVG approach when it comes to layout}