\chapter{Introduction}
\label{chap:Introduction}

The web is an indispensable part of modern society, and with the
ever-growing amounts of available data, visualizations which
effectively communicate this data are an essential part of it.
Motivated by the increasing variety of devices used to access the web,
responsive design has been established as one of the core pillars of
designing web content to ensure that it is easily accessible to users
regardless of the characteristics of their devices. Even though most
web content is already designed responsively, charts and
visualizations are often only embedded in a static form, which does
not adapt or only minimally adapts to different device
characteristics. The academic field of responsive visualization is
still in its early stages, but some research from various authors
regarding scalable visualizations
\parencite{BuildingRespDataVisForTheWeb,LearningRespDataVis}, design
patterns
\parencite{RespVisTalk,RespVisPage,RespVis,TechniquesForFlexibleRespVisDesign,
  DesignPatternsTradeOffsRespVis}, and tools
\parencite{TechniquesForFlexibleRespVisDesign} has already been done
and continues to emerge.


Many different software libraries to create visualizations for the web
exist, but they all have their shortcomings regarding usability,
extensibility, and responsiveness. One of the most popular libraries
for creating visualizations is D3 \parencite{D3,D3JS}, which provides
a low-level API to transform HTML and SVG documents based on
data. This document-based approach is rather powerful, as it allows
users to create whatever visualizations they wish, without being
hindered by the limitations of a custom renderer. However, building
visualizations by manually setting up their entire structure and
behavior can be tedious and requires deep knowledge of JavaScript
programming, D3, and the underlying rendering standard SVG.
%
Other visualization libraries like Vega \parencite{Vega} and its
extensions such as Cicero \parencite{Cicero}, which extends Vega
specifications with responsive transformations, focus on a
grammar-based approach to rendering visualizations. Visualization
grammars are very expressive and allow users to focus on a
visualization's high-level specification, but they tend to be rather
complex and are not always easy to understand. Furthermore, since the
actual rendering is abstracted away, visualization authors are limited
to the capabilities offered by a library's high-level API, which can
lead to restricted configurability.
%
Yet other types of visualization libraries are based on template-based
configuration, meaning that visualization authors only need to provide
data in a predefined format and the library then renders predefined
visualizations using this data. These template-based visualization
libraries are usually easy to use, but as with grammar-based
libraries, it can be hard to extend visualizations beyond the intended
configurations.


This thesis presents RespVis, a new software library for creating
presentational information visualizations for the web with a strong
emphasis on responsiveness. RespVis is an open-source library
\parencite{RespVisGitHub} written in TypeScript, which is designed as
an extension of D3 and focuses on rendering visualizations as
composite SVG documents. RespVis relies on CSS to style and position a
visualization's content, its JavaScript API has intentionally been
kept as minimal as possible. Visualization authors can use CSS media
queries in combination with Flexbox and Grid to position chart
components (such as title, axes, legend, and the chart itself)
responsively.

The main contribution of this work lies in a custom layouter which
uses the browser's own layout engine to enable the positioning of
SVG-based visualization components, which would otherwise be
unaffected by CSS layouting. Allowing visualization authors to
configure the layout of visualization components with CSS layouting
mechanisms like Flexbox and Grid leads to better responsive
capabilities than merely allowing them to style components with CSS.
Additionally, the use of CSS for styling and positioning allows the
application of other tools frequently used for responsive design like
media queries, and it also means that styles can easily be configured
and overwritten via the CSS cascade. Since RespVis renders and
configures visualizations using standardized web technologies such as
HTML, SVG, and CSS, visualization authors can work with technologies
all web developers are already familiar with and do not have to learn
a complex domain-specific language. Furthermore, the focus on standard
web technologies means that visualizations can easily be extended
beyond foreseen use cases, and it is less likely that visualization
authors are limited by restrictions of the library's API.



The first part of this thesis, in Chapters~\ref{chap:WebTechnologies}
to \ref{chap:ResponsiveInformationVisualization}, offers a broad view
of the technologies and related work into which this work is
embedded. Chapter~\ref{chap:WebTechnologies} introduces the various
web technologies upon which RespVis is built, the different ways of
embedding graphics into web pages, and the different layout engines
which were considered for laying out SVG chart elements.
Chapter~\ref{chap:InfoVis} gives an overview of the field of
information visualization and its history. Furthermore, some of the
more popular software libraries like D3, grammar-based libraries like
Vega, and template-based libraries like Highcharts used to create
information visualizations for the web are examined and compared
regarding their capabilities to make visualizations responsive.
Chapter~\ref{chap:ResponsiveInformationVisualization} looks in more
detail at the research around responsive visualizations.
Specifically, the topic of responsive patterns is introduced, and
their application is demonstrated with concrete examples from both
academic and other sources.

The second part of this thesis, in Chapters~\ref{chap:RespVis} to
\ref{chap:Outlook}, discusses the technical details of RespVis.
Chapter~\ref{chap:RespVis} introduces the library, its design pillars,
naming conventions, and project setup. Chapter~\ref{chap:Modules}
describes the implementation of RespVis by examining the various
packages and modules of the library. This chapter also discusses the
implementation and implications of the custom layouter which enables
layout of the elements of SVG-based visualizations using CSS layout
mechanisms. Chapter~\ref{chap:Usage} demonstrates the usage of
RespVis' modules to create responsive visualizations and explains how
different responsive patterns can be implemented. Finally,
Chapter~\ref{chap:Outlook} looks at potential future work to be done
on the library.

